---
title: "Feb20_Gender_Test"
author: "Jinlin He, Yuan Wang"
date: "February 20, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description
* ####Goal
We want to test within selected sample that if, when consider gender only, female and male use different time to finish the game. The overall goal is to see if students in different groups with sample size less than 50 will achieve different conclusions.

* ####Hypothesis
We hypothesis that gender influences TimeUsed.

* ####Method
    + We first filter out a gender data using v1label1.
    + Then, we mutate a column to the gender data where if the user was male we denote as 1, and if female we denote as 0.
    + We use a for loop to select all groupID's whose sample size is less then 50 and greater than 5. Also we use only groups that have more than 1 male and 1 female observations. (Otherwise t.test would fail).
    + We perform the two-sample t-test within each group.
    + We then perform a two-sample t-test for the over all gender data and yield the p-value as a comparison
  

## Libraries

```{r, include=FALSE}
library(plyr)
library(dplyr)
library(rmarkdown)
library(rprojroot)
library(ggplot2)
library(mosaic)
library(data.table)
library(xlsx)
```

## Import and Trim Data

We use only untimed data with non-zero time used recording.

```{r}
Original <- read.csv("Original.csv")
shape <- Original
shape_untime <- shape[shape$requestedTime==0,]
shape_untimed <- shape_untime[shape_untime$timeUsed!=0,]
```

We also changed some numerical variables to factors since they have only a fixed number of levels. We also changed the unit of the time data from milliseconds to seconds.

```{r}
shape_untimed$numShapes <- as.factor(shape_untimed$numShapes)
shape_untimed$matchingScheme <- as.factor(shape_untimed$matchingScheme)
shape_untimed$requestedTime <- as.factor(shape_untimed$requestedTime)
shape_untimed$timeUsed <- as.numeric(shape_untimed$timeUsed)
shape_untimed$timerDisplay <- as.factor(shape_untimed$timerDisplay)
shape_untimed$numErrors <- as.numeric(shape_untimed$numErrors)

shape_untimed <- mutate(shape_untimed, TimeUsedSec = shape_untimed$timeUsed/1000)
```

We further trimed the data to include a varible column named Gender. 

```{r}
gender <- filter(shape_untimed, tolower(strtrim(shape_untimed$v1label,3))=="gen" |
                   tolower(strtrim(shape_untimed$v1label,3))=="sex" |
                   tolower(strtrim(shape_untimed$v1label,4))=="male" |
                   tolower(strtrim(shape_untimed$v1label,1))=="f")

case <- (tolower(strtrim(gender$v1value,1)) == "m") & (tolower(strtrim(gender$v2value,1)) =="f")
gender1 <- gender[(strtrim(gender$v1value,1) != "1") &
                    (gender$v1value != "2") &
                    (gender$v1label != "Gender\rOrder") &
                    (gender$v1label != "female\rorder") &
                    (gender$v1value != "0") &
                    (tolower(gender$v1value) != "morf") &
                    (gender$studentID != "mb") &
                    (gender$studentID != "31207") &
                    (gender$v1value != "cat") &
                    (gender$studentID != "3659") &
                    (tolower(gender$v1value) != "attempt") &
                    (!case),]

gender1 <- mutate(gender1, 
                  gender = as.factor(ifelse(
                    pmax((tolower(strtrim(gender1$v1label,1)) == "m"),
                         (tolower(strtrim(gender1$v1value,1)) == "m")), 
                    1, 
                    ifelse(pmax(tolower(strtrim(gender1$v1label,1)) == "f",
                                (tolower(strtrim(gender1$v1value,1)) =="f")), 
                           0, 
                           "Inconclusive"))))

```

## Extract GroupID Information

Then, within the gender data, we identified the group names with sample size between 5 and 50. We make a vector of groupID of our desired groups.

```{r}
tb <- as.data.frame(table(gender1$groupID))
colnames(tb)[1]<-"groupID" ##Setting variable names for clarity

##Cross Tabulation to get sample size info for each group. 
tb1 <-as.data.frame(table(gender1$groupID,gender1$gender)) 
colnames(tb1)[1]<- "groupID"
colnames(tb1)[2]<- "gender"

##Groups to be removed
remove_gp <- c()
for(i in 1:length(tb1$Freq)){
  if(tb1$Freq[i] <= 1) {
    cbind(remove_gp,tb1$groupID)
  }
}
##Removing groups with sample size for either M/F factor level less than 1
group_ID <- as.character(tb$groupID)
for (i in 1:length(groupID)) {
  if (groupID[i]==)
      
}
##Select groups that are under size 50
tb2 <- tb[tb$Freq >= 5 & tb$Freq <= 50,]

##Create a vector of groupID's whose size is between 25 and 50
selected_groupID <- as.character(tb2$Var1)
```


## T-test and Boxplots

We will write a function to conduct a two-sided two sample t-test. We use the Student's t.test.
Our Null Hypothesis is that consider only gender, female and male use same time.
Our Alternative Hypothesis is that female and male use different times.
Since t.test does not work when either sample has less than 2 observations, we further filter the groupID so that each group has at least two female and two male observations.
We will show the boxplot of all groups using gender as the independet variable and time used (seconds) as the dependent variable. In the title of each graph, we will specify individual sample size and p-value. Choose alpha = 0.1, we color the those plots whose group reject the null hypothesis of no difference in green.

```{r, echo=TRUE}
par(mar=c(1,1,1,1))
par(mfrow = c(4,5))
groupName <- c()
pvalues <- c()
for (i in 1:length(selected_groupID)) {
  female <- gender2[gender2$groupID == selected_groupID[i] & gender2$gender == 0,]$TimeUsedSec
  male <- gender2[gender2$groupID == selected_groupID[i] & gender2$gender == 1,]$TimeUsedSec
  if (length(female) > 1 & length(male) > 1) {
    groupName <- cbind(groupName, selected_groupID[i])
    p <- round(t.test(female, male)$p.value, digits = 3)
    pvalues <- cbind(pvalues, p)
    data1 <- gender2[gender2$groupID == selected_groupID[i],]
    if (p < 0.1) {
      if (mean(female) < mean(male)) {
      boxplot(TimeUsedSec ~ gender,data=data1, col="green", main=paste("n=", dim(data1)[1], ", pval=", p), xlab="Gender (1=Male, 0=Female)", ylab="Time Used Seconds")
    } else {
      boxplot(TimeUsedSec ~ gender,data=data1, col="red", main=paste("n=", dim(data1)[1], ", pval=", p), xlab="Gender (1=Male, 0=Female)", ylab="Time Used Seconds")
    }
  } else {
    boxplot(TimeUsedSec ~ gender,data=data1, main=paste("n=", dim(data1)[1], ", pval=", p), xlab="Gender (1=Male, 0=Female)", ylab="Time Used Seconds")
  }
    
  write.csv(gender2[gender2$groupID == selected_groupID[i],], file =paste(selected_groupID[i], "group.csv"))
  }
}
```





We can also draw a boxplot for the over gender2 data and calculate a p-value as comparison to the group-wise plots and p-values.



```{r}
temp <- list.files(pattern = '*group.csv')
data <- rbind(read.csv(temp[1])[-c(1)], read.csv(temp[2])[-c(1)])
data <- rbind(data, read.csv(temp[3])[-c(1)])
data <- rbind(data, read.csv(temp[4])[-c(1)])
data <- rbind(data, read.csv(temp[5])[-c(1)])
data <- rbind(data, read.csv(temp[6])[-c(1)])              
data <- rbind(data, read.csv(temp[7])[-c(1)])
data <- rbind(data, read.csv(temp[8])[-c(1)])
data <- rbind(data, read.csv(temp[9])[-c(1)])
data <- rbind(data, read.csv(temp[10])[-c(1)])
data <- rbind(data, read.csv(temp[11])[-c(1)])
data <- rbind(data, read.csv(temp[12])[-c(1)])
data <- rbind(data, read.csv(temp[13])[-c(1)])
data <- rbind(data, read.csv(temp[14])[-c(1)])
data <- rbind(data, read.csv(temp[15])[-c(1)])
data <- rbind(data, read.csv(temp[16])[-c(1)])
data <- rbind(data, read.csv(temp[17])[-c(1)])
data <- rbind(data, read.csv(temp[18])[-c(1)])

female <- data[data$gender == 0,]$TimeUsedSec
male <- data[data$gender == 1,]$TimeUsedSec
pvl2 <- round(t.test(female, male)$p.value, digits = 3)
boxplot(TimeUsedSec ~ gender,data=data, main=paste("n=", dim(data)[1], ", two sided pval=", pvl2), xlab="Gender (1=Male, 0=Female)", ylab="Time Used Seconds")
```
